% !TeX encoding = UTF-8
% !TeX spellcheck = sk_SK
% !TeX root=teoriaprocesovtakac.tex
\section{Kyslíkový konvertor}

Kyslíkový konvertor
- čo to je (o aký technologický proces sa jedná, vstupy, výstupy ...)
- rozdelenie na elementárne procesy
- stručne popísať jednotlivé elementárne procesy



The first commercial operation of steelmaking with oxygen top blowing in
the converter was in the early 19508 at Linz and Donawitz (Austria). This
manner of steelmaking became known as Linz-Donawitz or LD process.
For many years now, most of the steel has been made by top oxygen
blowing for which different names are given. For example, in European
steel plants the process is still called LD; in the UK, BOS (basic oxygen
steelmaking); in the Far East and America, BOF (basic oxygen furnace), with
the exception of U.S. Steel where it is called BOP (basic oxygen process).

In the early 1970s, a bottom-blown oxygen steelmaking process was
developed in Canada and Germany. This process, known as OBM in Europe and Q-BOP elsewhere, was in full size commercial operation by the
mid 19708 in U.S. Steel plants followed by several plants in Europe and
Japan. The tuyeres, mounted in a removable bottom, consist of a central
pipe for blowing oxygen together with burnt lime, and an annular gap
around the central pipe for the passage of gaseous hydrocarbon, e.g. propane or natural gas (CH4). Upon contact with liquid steel the hydrocarbon
dissociates to C and H2 with the absorption of heat. This endothermic
reaction suppresses overheating of the tuyere tip by the exothermic reaction of oxygen with liquid steel.

Further developments in oxygen steelmaking led to the present practices of various types of top and bottom blowing known as combined
blowing, as illustrated schematically in Fig. 8.1 for BOF and Q-BOP processes. There is also post combustion of CO in the upper part of the vessel
to generate additional heat for steelmaking.

The presentation in this chapter is on the fundamentals of steelmaking
reactions and assessment of the state of slag-metal reactions at the end of
oxygen blowing, i.e. at first turndown of the vessel for sample taking. For
a description of steelmaking facilities and plant layout, reference may be
made to the U.S. Steel publication The Making, Shaping and Treating of Steel,
1985 edition.


-----------

Basic oxygen furnace (BOF) steelmaking is a complex process and dynamic model is very important for endpoint control. It is usually difficult to build a precise BOF endpoint dynamic model because many input variables affect the endpoint carbon content and temperature.


BOF is a widely preferred and effective steelmaking method
due to its high productivity and considerably low production cost. Therefore, almost 65\% of the total crude steel productions in the world are melted by using the BOF method. BOF steelmaking is a very complex chemical physical process. The quality of scrap iron changes from batch to batch. The grades of steel produced vary frequently, and the components of raw materials fluctuate largely.

The main objective of controlling oxygen converter steelmaking is to obtain prescribed parameters for the steel when it is tapped from the furnace, including weight, temperature, and each element content. In practical steelmaking process, the criterion whether the molten steel is acceptable or not is often decided by the endpoint carbon content and temperature.

-----------------------

Pri tejto metóde musí byť ale ešte dodávaná tavenina, aby dávka nevychladla.
Hotová oceľ sa potom vyleje z konvertora do panvy na ďalšie spracovanie.
Cenovo skoro najvýhodnejší spôsob výroby ocele pre veľké množstvá.
